---
title: "An odd relation: National Equity Prices and GDP Growth"
author: "Limberg, Daniel & Mueller, Lukas"
date: "April 2016"
output: pdf_document
---

```{r, echo=FALSE, message=FALSE, include=FALSE, cache=FALSE}
# Descriptive & Inferential Statistics
library(stargazer)
library(Rmisc)
library(ggplot2)
library(magrittr)
library(grid)


# Set Working Directory
try(setwd("/Users/Lukas/Documents/Git/PairAssignment3/"),silent=TRUE)
try(setwd("C:/Users/Dani/Documents/GitHub2/PairAssignment3/"),silent=TRUE)
getwd()


# Dynamical Link to first R script file
source("PairAssignment3_Code/PairAssignment3_Variables_LagsPercChanges.R")
```

# To Do

Mention Var?
Control for S&P 500?
Movement of Global markets?
Role of market sentiments?

# Introduction

Focusing on four OECD countries (Germany, France, Great Britain and Japan) we investigate on the question whether there are different levels of statistical significance for the effect of GDP growth on equity prices. For equity prices, we take the four major national stock market indices of the countries of interest as our dependent variable: The DAX (Germany), CAC (France), FTSE (Great Britain) and the NIKKEi (Japan). We control for a row of other explanatory variables.
[Br]
We take the quarterly GDP growth rate of the USA as a proxy for the world economy. The economy of the USA has a prfound impact on other national business cycles, and national recessions in the US have sever repercussions across the globe. This was demonstrated again after 2007 when the US subprime mortgage crisis turned into a global recession. Moreover, the West Texas Intermediate and Brent Crude oil price changes are taken as an indicator of global economic health.

# Research Question and Hypothesis

Are there different levels of statistical significance for the effect of a nation's GDP growth on national equity prices?

H~0~: There are no differences regarding statistical significance for the effect of a nation's GDP on national equity prices. 

H~1~: There are differences regarding statistical significance for the effect of a nation's GDP on national equity prices.

# Descriptive Statistics

## Variables of Interest

We investigate on four different dependent variables: National stock indices. We obtain data from their average closing values.

**Dependent Variables**   | Official Title                      | Measure
 -------------------- | ----------------------------------- | -----------------------
 DAX (Germany)        | Deutscher Aktienindex               | 30 major Companies
 CAC (France)         | Cotation Assisté en Continu 40      | 40 highest Market Caps
 FTSE (Great Britain) | Fin. Times Stock Exchange 100 Index | 100 Companies
 NIKKEI (Japan)       | NIKKEI 225                          | 225 Equities

We control for various covariates in our model. 

* ECB Main Refinancing Operations (% change to previous rate)
* ECB Deposit Facility (% change to previous rate)
* GDP Growth USA (% change to previous quarter)
* West Texas Intermediate (WTI)
* Brent Crude

We also control for country-specific covariates.

* GDP Growth (% change to previous quarter)
* Unemployment
* Private Consumption


## Dependent Variable (Equity Prices): Plotting the course over time

```{r, echo=FALSE, message=FALSE, include=FALSE, cache=FALSE}
### Germany
p1 <- ggplot(data = merge10, aes(x = Date, y = DAX.Close)) + geom_line() + scale_x_continuous(breaks=seq(1991.1, 2015.4, 10)) + ggtitle("Germany | DAX") + theme(plot.title = element_text(size=10)) + labs(x = "Quarters", y = "Av. Closing Value") + theme(legend.position="none") + theme(axis.text=element_text(size=8),
        axis.title=element_text(size=8))
### Japan
p2 <- ggplot(data = merge10, aes(x = Date, y = NIK.Close)) + geom_line() + scale_x_continuous(breaks=seq(1991.1, 2015.4, 10)) + ggtitle("Japan | NIKKEI") + theme(plot.title = element_text(size=10)) + labs(x = "Quarters", y = "Av. Closing Value") + theme(legend.position="none") + theme(axis.text=element_text(size=8),
        axis.title=element_text(size=8))
### France
p3 <- ggplot(data = merge10, aes(x = Date, y = CAC.Close)) + geom_line() + scale_x_continuous(breaks=seq(1991.1, 2015.4, 10)) + ggtitle("France | CAC40") + theme(plot.title = element_text(size=10)) + labs(x = "Quarters", y = "Av. Closing Value") + theme(legend.position="none") + theme(axis.text=element_text(size=8),
        axis.title=element_text(size=8))
### Great Britain
p4 <- ggplot(data = merge10, aes(x = Date, y = FTSE.Close)) + geom_line() + scale_x_continuous(breaks=seq(1991.1, 2015.4, 10)) + ggtitle("Great Britain | FTSE") + theme(plot.title = element_text(size=10)) + labs(x = "Quarters", y = "Av. Closing Value") + theme(legend.position="none") + theme(axis.text=element_text(size=8),
        axis.title=element_text(size=8))
```

```{r, echo=FALSE}
multiplot(p1, p2, p3, p4, cols = 2)
```


## Key Independent Variable: Quarterly GDP growth 

```{r, echo=FALSE, message=FALSE, include=FALSE, cache=FALSE}
farben <- c("firebrick1", "mediumspringgreen", "deepskyblue2", "chocolate4", "cyan2")
### Germany
p5 <- ggplot(data = merge10, aes(x = Date, y = DEU.GDP)) + geom_line() + scale_x_continuous(breaks=seq(1991.1, 2015.4, 10)) + ggtitle("Germany") + theme(plot.title = element_text(size=10)) + labs(x = "Quarters", y = "% Change") + theme(axis.text=element_text(size=8),
        axis.title=element_text(size=8)) + theme(legend.position="none")
### Japan
p6 <- ggplot(data = merge10, aes(x = Date, y = JPN.GDP)) + geom_line() + scale_x_continuous(breaks=seq(1991.1, 2015.4, 10)) + ggtitle("Japan") + theme(plot.title = element_text(size=10)) + labs(x = "Quarters", y = "% Change") + theme(axis.text=element_text(size=8),
        axis.title=element_text(size=8)) + theme(legend.position="none")
### France
p7 <- ggplot(data = merge10, aes(x = Date, y = FRA.GDP)) + geom_line() + scale_x_continuous(breaks=seq(1991.1, 2015.4, 10)) + ggtitle("France") + theme(plot.title = element_text(size=10)) + labs(x = "Quarters", y = "% Change") + theme(axis.text=element_text(size=8),
        axis.title=element_text(size=8)) + theme(legend.position="none")
### Great Britain
p8 <- ggplot(data = merge10, aes(x = Date, y = GBR.GDP)) + geom_line() + scale_x_continuous(breaks=seq(1991.1, 2015.4, 10)) + ggtitle("Great Britain") + theme(plot.title = element_text(size=10)) + labs(x = "Quarters", y = "% Change") + theme(axis.text=element_text(size=8),
        axis.title=element_text(size=8)) + theme(legend.position="none")
### USA
p9 <- ggplot(data = merge10, aes(x = Date, y = USA.GDP)) + geom_line() + scale_x_continuous(breaks=seq(1991.1, 2015.4, 10)) + ggtitle("USA") + theme(plot.title = element_text(size=10)) + labs(x = "Quarters", y = "% Change") + theme(axis.text=element_text(size=8),
        axis.title=element_text(size=8)) + theme(legend.position="none")
```

```{r, echo=FALSE}
multiplot(p5, p6, p7, p8, p9, cols = 2)
```

## Measures of Central Tendency for Dependent Variables

```{r, echo=FALSE}
### N, Mean St. Deviations, Min and Max for all Variables
stargazer(merge10, type = "text", title="Descriptive statistics", digits=1)
```

### Median

```{r, echo=FALSE}
for (i in 3:6) {
  merge10[, i] %>% 
    median() %>%
    paste(names(merge10)[i], ., "\n") %>%
    cat()
}
```

## Distribution of Dependent Variables: Average Closing Values

```{r, echo=FALSE}
par(mfrow=c(2,2))
hist(merge10$DAX.Close,
     main="Germany | DAX", 
     col="black", 
     breaks = 15,
     xlab = "Average Closing Values",
     ylab = "Frequency")
hist(merge10$CAC.Close,
     main="France | CAC", 
     col="black", 
     breaks = 15,
     xlab = "Average Closing Values",
     ylab = "Frequency")
hist(merge10$NIK.Close,
     main="Japan | Nikkei", 
     col="black", 
     breaks = 15,
     xlab = "Average Closing Values",
     ylab = "Frequency")
hist(merge10$FTSE.Close,
     main="Great Britain | FTSE", 
     col="black", 
     breaks = 15,
     xlab = "Average Closing Values",
     ylab = "Frequency")
```

## Joint Distributions of DV and Key IV (no lag)

```{r, echo=FALSE, message=FALSE, include=FALSE, cache=FALSE}
##DEU
p15 <- ggplot(merge10, aes(x = DEU.GDP, y = DAX.Close)) + 
  geom_point() + 
  geom_smooth() +
  ggtitle("Germany") +
  xlab("GDP Growth") +
  ylab("DAX Av. Closing Value") +
  theme(axis.text=element_text(size=8),
        axis.title=element_text(size=8)) +
  theme(plot.title = element_text(size=10)) 

## JPN
p16 <- ggplot(merge10, aes(x = JPN.GDP, y = NIK.Close)) + 
  geom_point() + 
  geom_smooth() +
  ggtitle("Japan") +
  xlab("GDP Growth") +
  ylab("Nikkei Av. Closing Value") +
  theme(axis.text=element_text(size=8),
        axis.title=element_text(size=8)) +
  theme(plot.title = element_text(size=10)) 

## FRA
p17 <- ggplot(merge10, aes(x = FRA.GDP, y = CAC.Close)) + 
  geom_point() + 
  geom_smooth() +
  ggtitle("France") +
  xlab("GDP Growth") +
  ylab("CAC Av. Closing Value") +
  theme(axis.text=element_text(size=8),
        axis.title=element_text(size=8)) +
  theme(plot.title = element_text(size=10)) 

## GBR
p18 <- ggplot(merge10, aes(x = GBR.GDP, y = FTSE.Close)) + 
  geom_point() + 
  geom_smooth() +
  ggtitle("Great Britain") +
  xlab("GDP Growth") +
  ylab("FTSE Av. Closing Value") +
  theme(axis.text=element_text(size=8),
        axis.title=element_text(size=8)) +
  theme(plot.title = element_text(size=10)) 
```

```{r, echo=FALSE}
multiplot(p15, p16, p17, p18, cols = 2)
```

# Inferential Statistics

## Basic Model: Pooled OLS

```{r, echo=FALSE, message=FALSE, include=FALSE, cache=FALSE}
summary(M1 <- lm(DAX.Close ~ DEU.GDP, data = merge10))
summary(M2 <- lm(NIK.Close ~ JPN.GDP, data = merge10))
summary(M3 <- lm(CAC.Close ~ FRA.GDP, data = merge10))
summary(M4 <- lm(FTSE.Close ~ GBR.GDP, data = merge10))
```

```{r, echo=FALSE}
stargazer(M1, M2, M3, M4, type="text", dep.var.labels=c("DAX","Nikkei", "CAC", "FTSE"),
          covariate.labels=c("GDP Growth"))
```

The above models show us the realtionship between national GDP and the respective stock indices. All coefficient have a positive sign but are statistically insignificant at the 5 per cent level which suggest that there is no effect differnet from zero. Only the coefficient for French GDP growth is significant at the 10 per cent level. The y-intercepts for all four regressions are positive and statistically sgnificant at the 1 per cent level.

### Heteroscedasticity Diagnose 

```{r, echo=FALSE}
par(mfrow=c(2,2))
plot(M1, which = 1)
  title("Germany")
plot(M2, which = 1)
  title("Japan")
plot(M3, which = 1)
  title("France")
plot(M4, which = 1)
  title("Britain")
```

Looking at the four graphs we see that the basic models suffer to a certain degree from heteroskedasticity. This stems from outliers which bias the results in the graphs. We conclude that the overall degree of hetereoskedasticity is acceptable since the general relationship seems to be random.

### Diagnose of non-normality of Errors

```{r, echo=FALSE}
par(mfrow=c(2,2))
plot(M1, which = 2)
  title("Germany")
plot(M2, which = 2)
  title("Japan")
plot(M3, which = 2)
  title("France")
plot(M4, which = 2)
  title("Britain")
```

The graphs give and idea of the distribution of the error term in the regression

## Expanded Model: Pooled OLS

```{r, echo=FALSE, message=FALSE, include=FALSE, cache=FALSE}
summary(M5 <- lm(DAX.Close ~ DEU.GDP + L.DEU.unempl + L.DEU.prvconsm + L.ECB.MRO.change + L.ECB.dep.change + L.WTI.dollar.change + L.Brent.dollar.change, data = merge10))
  
summary(M6 <- lm(NIK.Close ~ JPN.GDP + L.JPN.unempl + L.JPN.prvconsm + L.ECB.MRO.change + L.ECB.dep.change + L.WTI.dollar.change + L.Brent.dollar.change, data = merge10))

summary(M7 <- lm(CAC.Close ~ FRA.GDP + L.FRA.unempl + L.FRA.prvconsm + L.ECB.MRO.change + L.ECB.dep.change + L.WTI.dollar.change + L.Brent.dollar.change, data = merge10))

summary(M8 <- lm(FTSE.Close ~ GBR.GDP + L.GBR.unempl + L.GBR.prvconsm + L.ECB.MRO.change + L.ECB.dep.change + L.WTI.dollar.change + L.Brent.dollar.change, data = merge10))
```

```{r, echo=FALSE}
stargazer(M5, M6, M7, M8, type="text", dep.var.labels=c("DAX","Nikkei", "CAC", "FTSE"),
          covariate.labels=c("GDP Growth", "Unemployment lagged", "Consumption Spending lagged", "ECB MRO lagged", "ECB Depos. Facility lagged"))
```

## Heteroscedasticity Diagnose

```{r, echo=FALSE}
par(mfrow=c(2,2))
plot(M5, which = 1)
title("Germany")
plot(M6, which = 1)
title("Japan")
plot(M7, which = 1)
title("France")
plot(M8, which = 1)
title("Britain")
```

## Diagnose of non-normality of Errors

```{r, echo=FALSE}
par(mfrow=c(2,2))
plot(M5, which = 2)
title("Germany")
plot(M6, which = 2)
title("Japan")
plot(M7, which = 2)
title("France")
plot(M8, which = 2)
title("Britain")
```

# Expanded Model: Pooled OLS w/ year dummies

```{r, echo=FALSE, message=FALSE, include=FALSE, cache=FALSE}

summary(P1 <- lm(DAX.Close ~ DEU.GDP + L.DEU.unempl + L.DEU.prvconsm + L.ECB.MRO.change + L.ECB.dep.change  + L.WTI.dollar.change + L.Brent.dollar.change + year, data = merge10))

summary(P2 <- lm(NIK.Close ~ JPN.GDP + L.JPN.unempl + L.JPN.prvconsm + L.ECB.MRO.change + L.ECB.dep.change  + L.WTI.dollar.change + L.Brent.dollar.change + year, data = merge10))

summary(P3 <- lm(CAC.Close ~ FRA.GDP + L.FRA.unempl + L.FRA.prvconsm + L.ECB.MRO.change + L.ECB.dep.change  + L.WTI.dollar.change + L.Brent.dollar.change + year, data = merge10))

summary(P4 <- lm(FTSE.Close ~ GBR.GDP + L.GBR.unempl + L.GBR.prvconsm + L.ECB.MRO.change + L.ECB.dep.change  + L.WTI.dollar.change + L.Brent.dollar.change + year, data = merge10))
```

```{r, echo=FALSE}
stargazer(P1, P2, P3, P4, type="text", dep.var.labels=c("DAX","Nikkei", "CAC", "FTSE"),
          covariate.labels=c("GDP Growth", "Unemployment lagged", "Consumption Spending lagged", "ECB MRO lagged", "ECB Depos. Facility lagged", "WTI lagged", "Brent lagged", "Year"))
```

## Confidence Intervals

```{r, echo=FALSE}
confint(P1)
confint(P2)
confint(P3)
confint(P4)
```

## Heteroscedasticity Diagnose

```{r, echo=FALSE}
#par(mfrow=c(2,2))

#plot(P1, which = 1)
#title("Germany")

#plot(P2, which = 1)
#title("Japan")

#plot(P3, which = 1)
#title("France")

#plot(P4, which = 1)
#title("Britain")
```

## Diagnose of non-normality of Errors

```{r, echo=FALSE}
#par(mfrow=c(2,2))

#plot(P1, which = 2)
#title("Germany")

#plot(P2, which = 2)
#title("Japan")

#plot(P3, which = 2)
#title("France")

#plot(P4, which = 2)
#title("Britain")
```

# Inferential Statistics - US ecomomy

```{r, echo=FALSE, message=FALSE, include=FALSE, cache=FALSE}
M9 <- lm(DAX.Close ~ DEU.GDP + L.DEU.unempl + L.DEU.prvconsm + L.ECB.MRO.change + L.ECB.dep.change  + L.WTI.dollar.change + L.Brent.dollar.change + year + USA.GDP, data = merge10)

M10 <- lm(NIK.Close ~ JPN.GDP + L.JPN.unempl + L.JPN.prvconsm + L.ECB.MRO.change + L.ECB.dep.change  + L.WTI.dollar.change + L.Brent.dollar.change + year + USA.GDP, data = merge10)

M11 <- lm(CAC.Close ~ FRA.GDP + L.FRA.unempl + L.FRA.prvconsm + L.ECB.MRO.change + L.ECB.dep.change  + L.WTI.dollar.change + L.Brent.dollar.change + year + USA.GDP, data = merge10)

M12 <- lm(FTSE.Close ~ GBR.GDP + L.GBR.unempl + L.GBR.prvconsm + L.ECB.MRO.change + L.ECB.dep.change  + L.WTI.dollar.change + L.Brent.dollar.change + year + USA.GDP, data = merge10)
```

```{r, echo=FALSE}
stargazer(M9, M10, M11, M12, type="text", dep.var.labels=c("DAX","Nikkei", "CAC", "FTSE"),
covariate.labels=c("GDP Growth", "Unemployment lagged", "Consumption Spending lagged", "ECB MRO lagged", "ECB Depos. Facility lagged", "WTI lagged", "Brent lagged", "Year", "USA.GDP"))
```

## Confidence Intervals

```{r, echo=FALSE}
confint(M9)
confint(M10)
confint(M11)
confint(M12)
```

## Heteroscedasticity Diagnose

```{r, echo=FALSE}
#par(mfrow=c(2,2))
#plot(M9, which = 1)
#title("Germany")
#plot(M10, which = 1)
#title("Japan")
#plot(M11, which = 1)
#title("France")
#plot(M12, which = 1)
#title("Britain")
```

## Diagnose of non-normality of Errors

```{r, echo=FALSE}
#par(mfrow=c(2,2))
#plot(M9, which = 2)
#title("Germany")
#plot(M10, which = 2)
#title("Japan")
#plot(M11, which = 2)
#title("France")
#plot(M12, which = 2)
#title("Britain")
```
