---
title: "An odd relation: National Equity Prices and GDP Growth"
author: "Limberg, Daniel & Mueller, Lukas"
date: "April 2016"
output: pdf_document
---

```{r, echo=FALSE, message=FALSE, include=FALSE, cache=FALSE}
# Descriptive & Inferential Statistics
library(stargazer)
library(Rmisc)
library(ggplot2)
library(magrittr)
library(grid)
library(plm)


# Set Working Directory
try(setwd("/Users/Lukas/Documents/Git/PairAssignment3/"),silent=TRUE)
try(setwd("C:/Users/Dani/Documents/GitHub2/PairAssignment3/"),silent=TRUE)
getwd()


# Dynamical Link to first R script file
source("PairAssignment3_Code/PairAssignment3_Variables_LagsPercChanges.R")


### Descriptive Statistics ###

# Subset Data
data <- as.data.frame(panel)
USA <- subset(data, iso3c == "USA")
DEU <- subset(data, iso3c == "DEU")
GBR <- subset(data, iso3c == "GBR")
JPN <- subset(data, iso3c == "JPN")
FRA <- subset(data, iso3c == "FRA")
ALL <- subset(data, iso3c == "JPN" | iso3c == "DEU" | iso3c == "GBR" | iso3c == "FRA" | iso3c == "USA")
ALL <- ALL[ ,c(1,2,8,14,20,26,2:7,9:13,15:19,21:25,27:41)]
```

# Introduction



## Dependent Variable (Equity Prices): Plotting the course over time

```{r, echo=FALSE, message=FALSE, include=FALSE, cache=FALSE}
### Germany
p1 <- ggplot(data = DEU, aes(x = Date, y = DAX.Close, group = iso3c, color = iso3c)) + geom_line() + ggtitle("DAX") + theme(plot.title = element_text(size=9)) + labs(x = "Quarters", y = "Av. Closing Value")
p1 + scale_colour_discrete(name  ="Country", labels=c("Germany"))
### Japan
p2 <- ggplot(data = JPN, aes(x = Date, y = NIK.Close, group = iso3c, color = iso3c)) + geom_line() + ggtitle("NIKKEI") + theme(plot.title = element_text(size=9)) + labs(x = "Quarters", y = "Av. Closing Value")
p2 + scale_colour_discrete(name  ="Country", labels=c("Japan"))
### France
p3 <- ggplot(data = FRA, aes(x = Date, y = CAC.Close, group = iso3c, color = iso3c)) + geom_line() + ggtitle("CAC40") + theme(plot.title = element_text(size=9)) + labs(x = "Quarters", y = "Av. Closing Value")
p3 + scale_colour_discrete(name  ="Country", labels=c("France"))
### Great Britain
p4 <- ggplot(data = GBR, aes(x = Date, y = FTSE.Close, group = iso3c, color = iso3c)) + geom_line() + ggtitle("FTSE") + theme(plot.title = element_text(size=9)) + labs(x = "Quarters", y = "Av. Closing Value")
p4 + scale_colour_discrete(name  ="Country", labels=c("Britain"))
```

```{r, echo=FALSE}
multiplot(p1, p2, p3, p4, cols = 2)
```


## Key Independent Variable: Quarterly GDP growth 

```{r, echo=FALSE, message=FALSE, include=FALSE, cache=FALSE}
### Germany
p5 <- ggplot(data = DEU, aes(x = Date, y = GDPq.gr, group, group = iso3c, color = iso3c)) + geom_line() + ggtitle("Quarterly GDP growth") + theme(plot.title = element_text(size=9)) + labs(x = "Quarters", y = "% Change to prev. Quarter")
p5 + scale_colour_discrete(name  ="Country", labels=c("Germany"))
### Japan
p6 <- ggplot(data = JPN, aes(x = Date, y = GDPq.gr, group, group = iso3c, color = iso3c)) + geom_line() + ggtitle("Quarterly GDP growth") + theme(plot.title = element_text(size=9)) + labs(x = "Quarters", y = "% Change to prev. Quarter")
p6 + scale_colour_discrete(name  ="Country", labels=c("Japan"))
### France
p7 <- ggplot(data = FRA, aes(x = Date, y = GDPq.gr, group, group = iso3c, color = iso3c)) + geom_line() + ggtitle("Quarterly GDP growth") + theme(plot.title = element_text(size=9)) + labs(x = "Quarters", y = "% Change to prev. Quarter")
p7 + scale_colour_discrete(name  ="Country", labels=c("France"))
### Great Britain
p8 <- ggplot(data = GBR, aes(x = Date, y = GDPq.gr, group, group = iso3c, color = iso3c)) + geom_line() + ggtitle("Quarterly GDP growth") + theme(plot.title = element_text(size=9)) + labs(x = "Quarters", y = "% Change to prev. Quarter")
p8 + scale_colour_discrete(name  ="Country", labels=c("Britain"))
### USA
p9 <- ggplot(data = USA, aes(x = Date, y = GDPq.gr, group, group = iso3c, color = iso3c)) + geom_line() + ggtitle("Quarterly GDP growth") + theme(plot.title = element_text(size=9)) + labs(x = "Quarters", y = "% Change to prev. Quarter")
p9 + scale_colour_discrete(name  ="Country", labels=c("USA"))
### All growth rates
p10 <- ggplot(data = ALL, aes(x = Date, y = GDPq.gr, group = iso3c, color = iso3c)) + geom_line() + ggtitle("Quarterly GDP growth") + theme(plot.title = element_text(size=9)) + labs(x = "Quarters", y = "% Change to prev. Quarter")
p10 + scale_colour_discrete(name  ="Country", labels=c("Germany", "France", "Britain", "Japan", "USA"))
```

```{r, echo=FALSE}
multiplot(p5, p6, p7, p8, p9,p10, cols = 2)
```

# Measures of Central Tendency for Dependent Variables

## Mean

```{r, echo=FALSE}
for (i in 3:6) {
  ALL[, i] %>% 
    mean() %>%
    paste(names(ALL)[i], ., "\n") %>%
    cat()
}
```

## Median

```{r, echo=FALSE}
for (i in 3:6) {
  ALL[, i] %>% 
    median() %>%
    paste(names(ALL)[i], ., "\n") %>%
    cat()
}
```

## Standard Dev.

```{r, echo=FALSE}
for (i in 3:6) {
  ALL[, i] %>% #Error: could not find function "%>%"
    sd() %>%
    paste(names(ALL)[i], ., "\n") %>%
    cat()
}
```

# Distribution of Dependent Variables: Average Closing Values

```{r, echo=FALSE}
par(mfrow=c(2,2))
hist(DEU$DAX.Close,
     main="Germany | DAX", 
     col="black", 
     breaks = 15,
     xlab = "Average Closing Values",
     ylab = "Frequency")
hist(JPN$NIK.Close,
     main="Japan | Nikkei", 
     col="black", 
     breaks = 15,
     xlab = "Average Closing Values",
     ylab = "Frequency")
hist(FRA$CAC.Close,
     main="France | CAC", 
     col="black", 
     breaks = 15,
     xlab = "Average Closing Values",
     ylab = "Frequency")
hist(GBR$FTSE.Close,
     main="Great Britain | FTSE", 
     col="black", 
     breaks = 15,
     xlab = "Average Closing Values",
     ylab = "Frequency")
```

# Joint Distributions of DV and Key IV (no lag)

```{r, echo=FALSE, message=FALSE, include=FALSE, cache=FALSE}
##DEU
p15 <- ggplot(DEU, aes(x = GDPq.gr, y = DAX.Close)) + 
  geom_point() + 
  geom_smooth() +
  ggtitle("Germany") +
  xlab("GDP Growth") +
  ylab("DAX Av. Closing Value")
## JPN
p16 <- ggplot(JPN, aes(x = GDPq.gr, y = NIK.Close)) + 
  geom_point() + 
  geom_smooth() +
  ggtitle("Japan") +
  xlab("GDP Growth") +
  ylab("Nikkei Av. Closing Value")
## FRA
p17 <- ggplot(FRA, aes(x = GDPq.gr, y = CAC.Close)) + 
  geom_point() + 
  geom_smooth() +
  ggtitle("France") +
  xlab("GDP Growth") +
  ylab("CAC Av. Closing Value")
## GBR
p18 <- ggplot(GBR, aes(x = GDPq.gr, y = FTSE.Close)) + 
  geom_point() + 
  geom_smooth() +
  ggtitle("Great Britain") +
  xlab("GDP Growth") +
  ylab("FTSE Av. Closing Value")
```

```{r, echo=FALSE}
multiplot(p15, p16, p17, p18, cols = 2)
```

# Inferential Statistics

## Basic Model: OLS

```{r, echo=FALSE, message=FALSE, include=FALSE, cache=FALSE}
summary(M1 <- lm(DAX.Close ~ GDPq.gr, data = DEU))
summary(M2 <- lm(NIK.Close ~ GDPq.gr, data = JPN))
summary(M3 <- lm(CAC.Close ~ GDPq.gr, data = FRA))
summary(M4 <- lm(FTSE.Close ~ GDPq.gr, data = GBR))
```

```{r, echo=FALSE}
stargazer(M1, M2, M3, M4, type="text", dep.var.labels=c("DAX","Nikkei", "CAC", "FTSE"),
          covariate.labels=c("GDP Growth"))
```

### Heteroscedasticity Diagnose 

```{r, echo=FALSE}
par(mfrow=c(2,2))
plot(M1, which = 1)
  title("Germany")
plot(M2, which = 1)
  title("Japan")
plot(M3, which = 1)
  title("France")
plot(M4, which = 1)
  title("Britain")
```

### Diagnose of non-normality of Errors

```{r, echo=FALSE}
par(mfrow=c(2,2))
plot(M1, which = 2)
  title("Germany")
plot(M2, which = 2)
  title("Japan")
plot(M3, which = 2)
  title("France")
plot(M4, which = 2)
  title("Britain")
```

## Expanded Model: OLS

```{r, echo=FALSE, message=FALSE, include=FALSE, cache=FALSE}
summary(M5 <- lm(DAX.Close ~ GDPq.gr + L.unempl + L.cons.spend + L.ECB.MRO.ch + L.ECB.dep.ch + L.WTI + L.Brent, data = DEU))
summary(M6 <- lm(NIK.Close ~ GDPq.gr + L.unempl + L.cons.spend + L.ECB.MRO.ch + L.ECB.dep.ch + L.WTI + L.Brent, data = JPN))
summary(M7 <- lm(CAC.Close ~ GDPq.gr + L.unempl + L.cons.spend + L.ECB.MRO.ch + L.ECB.dep.ch + L.WTI + L.Brent, data = FRA))
summary(M8 <- lm(FTSE.Close ~ GDPq.gr + L.unempl + L.cons.spend + L.ECB.MRO.ch + L.ECB.dep.ch + L.WTI + L.Brent, data = GBR))
```

```{r, echo=FALSE}
stargazer(M5, M6, M7, M8, type="text", dep.var.labels=c("DAX","Nikkei", "CAC", "FTSE"),
          covariate.labels=c("GDP Growth", "Unemployment lagged", "Consumption Spending lagged", "ECB MRO lagged", "ECB Depos. Facility lagged"))
```

## Heteroscedasticity Diagnose

```{r, echo=FALSE}
par(mfrow=c(2,2))
plot(M5, which = 1)
title("Germany")
plot(M6, which = 1)
title("Japan")
plot(M7, which = 1)
title("France")
plot(M8, which = 1)
title("Britain")
```

## Diagnose of non-normality of Errors

```{r, echo=FALSE}
par(mfrow=c(2,2))
plot(M5, which = 2)
title("Germany")
plot(M6, which = 2)
title("Japan")
plot(M7, which = 2)
title("France")
plot(M8, which = 2)
title("Britain")
```

# Inferential Statistics - Panel Model: Pooled OLS

```{r, echo=FALSE, message=FALSE, include=FALSE, cache=FALSE}
panelall <- pdata.frame(ALL, index=c("iso3c", "Date"))

summary(P1 <- plm(DAX.Close ~ GDPq.gr + L.unempl + L.cons.spend + L.ECB.MRO.ch + L.ECB.dep.ch  + L.WTI + L.Brent, data = panelall, model = "pooling"))

summary(P2 <- plm(NIK.Close ~ GDPq.gr + L.unempl + L.cons.spend + L.ECB.MRO.ch + L.ECB.dep.ch  + L.WTI + L.Brent, data = panelall, model = "pooling"))

summary(P3 <- plm(CAC.Close ~ GDPq.gr + L.unempl + L.cons.spend + L.ECB.MRO.ch + L.ECB.dep.ch  + L.WTI + L.Brent, data = panelall, model = "pooling"))

summary(P4 <- plm(FTSE.Close ~ GDPq.gr + L.unempl + L.cons.spend + L.ECB.MRO.ch + L.ECB.dep.ch  + L.WTI + L.Brent, data = panelall, model = "pooling"))
```

```{r, echo=FALSE}
stargazer(P1, P2, P3, P4, type="text", dep.var.labels=c("DAX","Nikkei", "CAC", "FTSE"),
          covariate.labels=c("GDP Growth", "Unemployment lagged", "Consumption Spending lagged", "ECB MRO lagged", "ECB Depos. Facility lagged", "WTI lagged", "Brent lagged"))
```

## Confidence Intervals

```{r, echo=FALSE}
confint(P1)
confint(P2)
confint(P3)
confint(P4)
```

## Heteroscedasticity Diagnose

```{r, echo=FALSE}
#par(mfrow=c(2,2))

#plot(P1, which = 1)
#title("Germany")

#plot(P2, which = 1)
#title("Japan")

#plot(P3, which = 1)
#title("France")

#plot(P4, which = 1)
#title("Britain")
```

## Diagnose of non-normality of Errors

```{r, echo=FALSE}
#par(mfrow=c(2,2))

#plot(P1, which = 2)
#title("Germany")

#plot(P2, which = 2)
#title("Japan")

#plot(P3, which = 2)
#title("France")

#plot(P4, which = 2)
#title("Britain")
```

# Inferential Statistics - Panel Model: Fixed Effects

```{r, echo=FALSE, message=FALSE, include=FALSE, cache=FALSE}
M9 <- plm(DAX.Close ~ GDPq.gr + L.unempl + L.cons.spend + L.ECB.MRO.ch + L.ECB.dep.ch  + L.WTI + L.Brent, data = panelall, model = "within")

M10 <- plm(NIK.Close ~ GDPq.gr + L.unempl + L.cons.spend + L.ECB.MRO.ch + L.ECB.dep.ch  + L.WTI + L.Brent, data = panelall, model = "within")

M11 <- plm(CAC.Close ~ GDPq.gr + L.unempl + L.cons.spend + L.ECB.MRO.ch + L.ECB.dep.ch  + L.WTI + L.Brent, data = panelall, model = "within")

M12 <- plm(FTSE.Close ~ GDPq.gr + L.unempl + L.cons.spend + L.ECB.MRO.ch + L.ECB.dep.ch  + L.WTI + L.Brent, data = panelall, model = "within")
```

```{r, echo=FALSE}
stargazer(M9, M10, M11, M12, type="text", dep.var.labels=c("DAX","Nikkei", "CAC", "FTSE"),
covariate.labels=c("GDP Growth", "Unemployment lagged", "Consumption Spending lagged", "ECB MRO lagged", "ECB Depos. Facility lagged", "WTI lagged", "Brent lagged"))
```

## Confidence Intervals

```{r, echo=FALSE}
confint(M9)
confint(M10)
confint(M11)
confint(M12)
```

## Heteroscedasticity Diagnose

```{r, echo=FALSE}
#par(mfrow=c(2,2))
#plot(M9, which = 1)
#title("Germany")
#plot(M10, which = 1)
#title("Japan")
#plot(M11, which = 1)
#title("France")
#plot(M12, which = 1)
#title("Britain")
```

## Diagnose of non-normality of Errors

```{r, echo=FALSE}
#par(mfrow=c(2,2))
#plot(M9, which = 2)
#title("Germany")
#plot(M10, which = 2)
#title("Japan")
#plot(M11, which = 2)
#title("France")
#plot(M12, which = 2)
#title("Britain")
```
