---
title: "An odd relation: National Equity Prices and GDP Growth"
author: "Limberg, Daniel & Mueller, Lukas"
date: "April 2016"
output: pdf_document
---

```{r, echo=FALSE, message=FALSE, include=FALSE, cache=FALSE}
# Descriptive & Inferential Statistics
library(stargazer)
library(Rmisc)
library(ggplot2)
library(magrittr)
library(grid)
library(xtable)
library(car)

# Set Working Directory
try(setwd("/Users/Lukas/Documents/Git/PairAssignment3/"),silent=TRUE)
try(setwd("C:/Users/Dani/Documents/GitHub2/PairAssignment3/"),silent=TRUE)
getwd()


# Dynamical Link to first R script file
source("PairAssignment3_Code/PairAssignment3_Variables_LagsPercChanges.R")
```

# Introduction

In this assignment, we investigate on an odd relation: GDP growth and national equity prices; that is national stock indices. The intuitive assumption that one has automatically is fairly straight forward: The better the economy performs, the higher expectations of market participants, investments and trust in the future. In turn, we should - naturally - obtain an increase the level of national stock market indices. 

However, the more interesting question is beyond this intuitive assumption:
Focusing on four OECD countries (Germany, France, Great Britain and Japan) we investigate on the question whether there are different levels of statistical significance for the effect of GDP growth on equity prices. Note that we do not compare the size of the effects (i.e. coefficients) since we investigate on different dependent variables. Rather, we compare the statistical significance across the four countries of interest. For equity prices, we take the four major national stock market indices of the countries as our dependent variable: The DAX (Germany), CAC (France), FTSE (Great Britain) and the NIKKEI (Japan). We control for a row of other, lagged explanatory determinants - not only from within the national economy (e.g. unemployment rate) but also external, more global factors (e.g. oil prices). We elaborate on the specific covariates below. For our analysis we focus on the time period between the second quarter of 1999 (where ECB interest rates became relevant) and fourth quarter of 2015. 

# Research Question and Hypothesis

*Does national GDP growth lead to growing national stock market indices?*

* H~0~: There are statistically significant effects of national GDP growth on the growth rates of national stock market indices.

* H~1~: There are no statistically significant effects of national GDP growth on the growth rates of national stock market indices.

# Descriptive Statistics

## Variables of Interest

We investigate on four different dependent variables: The National stock indices of Germany, France, Great Britain and Japan. We obtain data from their average closing values.

**Dependent Variables**   | Official Title                      | Scope
 -------------------- | ----------------------------------- | -----------------------
 DAX (Germany)        | Deutscher Aktienindex               | 30 major Companies
 CAC (France)         | Cotation Assisté en Continu 40      | 40 highest Market Caps
 FTSE (Great Britain) | Fin. Times Stock Exchange 100 Index | 100 Companies
 NIKKEI (Japan)       | NIKKEI 225                          | 225 Equities

We control for various covariates in our model. We lag all our covariates by one quarter. 

* ECB Main Refinancing Operations (% change to previous rate)
* ECB Deposit Facility (% change to previous rate)
* GDP Growth USA (% change to previous quarter)
* West Texas Intermediate (WTI) (% change to previous quarter in U.S. $)
* Brent Crude (% change to previous quarter in U.S. $)

We take the quarterly GDP growth rate of the USA as a proxy for the world economy. The economy of the USA has a profound impact on other national business cycles, and national recessions in the US have sever repercussions across the globe. This was demonstrated again after 2007 when the US subprime mortgage crisis turned into a global recession. Moreover, the West Texas Intermediate and Brent Crude oil price changes are taken as an indicator of global economic health.

We also control for country-specific covariates.

* GDP Growth (% change to previous quarter)
* Unemployment (in %)
* Private Consumption (in % of GDP )

Note that for the descriptive statistics section all graphs apply the full range of our available data (i.e. Q1 1991 until Q4 2015). However, our actual analysis (under the inferential statistics section) only takes into account the time period between Q2 1999 and Q4 2015 since we also control for ECB interest rates. 
**Figure 1** presents a first glance at the four dependent variables from the first quarter in 1991 to the last quarter in 2015. Striking is the non-linearity of the graphs and the fact that the NIKKEI - quite contrary to the other indices - rather declined/stagnated. The DAX, CAC 40 and the FTSE shows fairly similar developments.  

## Figure 1: *Dependent Variables (Change of National Stock Indices) over time*

```{r, echo=FALSE, message=FALSE, include=FALSE, cache=FALSE}
### Germany
p1 <- ggplot(data = merge10, aes(x = Date, y = DAX.Close.change)) + geom_line() + scale_x_continuous(breaks=seq(1991.1, 2015.4, 10)) + ggtitle("Germany | DAX") + theme(plot.title = element_text(size=10)) + labs(x = "Quarters", y = "% change of Closing Value") + theme(legend.position="none") + theme(axis.text=element_text(size=8),
        axis.title=element_text(size=8))
### Japan
p2 <- ggplot(data = merge10, aes(x = Date, y = NIK.Close.change)) + geom_line() + scale_x_continuous(breaks=seq(1991.1, 2015.4, 10)) + ggtitle("Japan | NIKKEI") + theme(plot.title = element_text(size=10)) + labs(x = "Quarters", y = "% change of Closing Value") + theme(legend.position="none") + theme(axis.text=element_text(size=8),
        axis.title=element_text(size=8))
### France
p3 <- ggplot(data = merge10, aes(x = Date, y = CAC.Close.change)) + geom_line() + scale_x_continuous(breaks=seq(1991.1, 2015.4, 10)) + ggtitle("France | CAC40") + theme(plot.title = element_text(size=10)) + labs(x = "Quarters", y = "% change of Closing Value") + theme(legend.position="none") + theme(axis.text=element_text(size=8),
        axis.title=element_text(size=8))
### Great Britain
p4 <- ggplot(data = merge10, aes(x = Date, y = FTSE.Close.change)) + geom_line() + scale_x_continuous(breaks=seq(1991.1, 2015.4, 10)) + ggtitle("Great Britain | FTSE") + theme(plot.title = element_text(size=10)) + labs(x = "Quarters", y = "% change of Closing Value") + theme(legend.position="none") + theme(axis.text=element_text(size=8),
        axis.title=element_text(size=8))
```

```{r, echo=FALSE}
multiplot(p1, p2, p3, p4, cols = 2)
```

Another overview we should be interested in is given in **Figure 2**: The development of our key independent variables over the same time (i.e. Q1 1991 until Q4 2015). Note that here we also include the graph for the USA which is an external factor we assume to be influential on all four economies of interest (and in turn on the according national stock indices). Compared to the previous developments of the dependent variables, the key independent variables show a relatively more similar, congruent evolution. Remarkable is the major, well visible downturn in the aftermath of financial crisis in all five graphs.  

## Figure 2: *Key Independent Variables (Quarterly GDP growth to previous Quarter)* 

```{r, echo=FALSE, message=FALSE, include=FALSE, cache=FALSE}
farben <- c("firebrick1", "mediumspringgreen", "deepskyblue2", "chocolate4", "cyan2")
### Germany
p5 <- ggplot(data = merge10, aes(x = Date, y = DEU.GDP)) + geom_line() + scale_x_continuous(breaks=seq(1991.1, 2015.4, 10)) + ggtitle("Germany") + theme(plot.title = element_text(size=10)) + labs(x = "Quarters", y = "% Change") + theme(axis.text=element_text(size=8),
        axis.title=element_text(size=8)) + theme(legend.position="none")
### Japan
p6 <- ggplot(data = merge10, aes(x = Date, y = JPN.GDP)) + geom_line() + scale_x_continuous(breaks=seq(1991.1, 2015.4, 10)) + ggtitle("Japan") + theme(plot.title = element_text(size=10)) + labs(x = "Quarters", y = "% Change") + theme(axis.text=element_text(size=8),
        axis.title=element_text(size=8)) + theme(legend.position="none")
### France
p7 <- ggplot(data = merge10, aes(x = Date, y = FRA.GDP)) + geom_line() + scale_x_continuous(breaks=seq(1991.1, 2015.4, 10)) + ggtitle("France") + theme(plot.title = element_text(size=10)) + labs(x = "Quarters", y = "% Change") + theme(axis.text=element_text(size=8),
        axis.title=element_text(size=8)) + theme(legend.position="none")
### Great Britain
p8 <- ggplot(data = merge10, aes(x = Date, y = GBR.GDP)) + geom_line() + scale_x_continuous(breaks=seq(1991.1, 2015.4, 10)) + ggtitle("Great Britain") + theme(plot.title = element_text(size=10)) + labs(x = "Quarters", y = "% Change") + theme(axis.text=element_text(size=8),
        axis.title=element_text(size=8)) + theme(legend.position="none")
### USA
p9 <- ggplot(data = merge10, aes(x = Date, y = USA.GDP)) + geom_line() + scale_x_continuous(breaks=seq(1991.1, 2015.4, 10)) + ggtitle("USA") + theme(plot.title = element_text(size=10)) + labs(x = "Quarters", y = "% Change") + theme(axis.text=element_text(size=8),
        axis.title=element_text(size=8)) + theme(legend.position="none")
```

```{r, echo=FALSE}
multiplot(p5, p6, p7, p8, p9, cols = 2)
```

**Figure 3** summarizes the main characteristics for the depent and key indepent variables.

## Figure 3: *Measures of Central Tendency for DVs & Key IVs*

```{r, echo=FALSE, results="asis"}
### N, Mean St. Deviations, Min and Max for all Variables
stargazer(merge10[c("DAX.Close.change", "CAC.Close.change", "NIK.Close.change", "FTSE.Close.change", "L.DEU.GDP","L.FRA.GDP","L.JPN.GDP", "L.GBR.GDP", "L.USA.GDP")], 
          type = "latex",
          title = "Descriptive statistics",
          header = FALSE,
          digits=1)
```

The according Medians for the each dependent variable are as follows: 

```{r, echo=FALSE, results="asis"}
for (i in 65:68) {
  merge10[, i] %>% 
    median() %>%
    paste(names(merge10)[i], ., "\n") %>%
    cat()
}
```

In the next step, **Figure 4** summarizes the frequency distributions of the dependent variables. For now, we assume the variables to be normally distributed. In following research to this assignment we will consider log-transformation for the variables.

## Figure 4: *Distribution of Dependent Variables: Average Closing Values*

```{r, echo=FALSE}
par(mfrow=c(2,2))
hist(merge10$DAX.Close.change,
     main="Germany | DAX", 
     col="black", 
     breaks = 15,
     xlab = "Average Closing Values",
     ylab = "Frequency")
hist(merge10$CAC.Close.change,
     main="France | CAC", 
     col="black", 
     breaks = 15,
     xlab = "Average Closing Values",
     ylab = "Frequency")
hist(merge10$NIK.Close.change,
     main="Japan | Nikkei", 
     col="black", 
     breaks = 15,
     xlab = "Average Closing Values",
     ylab = "Frequency")
hist(merge10$FTSE.Close.change,
     main="Great Britain | FTSE", 
     col="black", 
     breaks = 15,
     xlab = "Average Closing Values",
     ylab = "Frequency")
```

Finally, we obtain the joint distributions between our dependent and key independent variables in **Figure 5**. Remarkable are the trends for a high GDP growth (i.e. over 2%) in the case of Germany and Japan. Here, the intuitive relation between higher GDP growth and a higher level of the index is violated. 

## Figure 5: *Joint Distributions of DV and Key IV*

```{r, echo=FALSE, message=FALSE, include=FALSE, cache=FALSE, warning=FALSE}
##DEU
p15 <- ggplot(merge10, aes(x = L.DEU.GDP, y = DAX.Close.change)) + 
  geom_point() + 
  geom_smooth() +
  ggtitle("Germany") +
  xlab("GDP Growth") +
  ylab("DAX Growth Rate in %") +
  theme(axis.text=element_text(size=8),
        axis.title=element_text(size=8)) +
  theme(plot.title = element_text(size=10)) 

## JPN
p16 <- ggplot(merge10, aes(x = L.JPN.GDP, y = NIK.Close.change)) + 
  geom_point() + 
  geom_smooth() +
  ggtitle("Japan") +
  xlab("GDP Growth") +
  ylab("Nikkei Growth Rate in %") +
  theme(axis.text=element_text(size=8),
        axis.title=element_text(size=8)) +
  theme(plot.title = element_text(size=10)) 

## FRA
p17 <- ggplot(merge10, aes(x = L.FRA.GDP, y = CAC.Close.change)) + 
  geom_point() + 
  geom_smooth() +
  ggtitle("France") +
  xlab("GDP Growth") +
  ylab("CAC Growth Rate in %") +
  theme(axis.text=element_text(size=8),
        axis.title=element_text(size=8)) +
  theme(plot.title = element_text(size=10)) 

## GBR
p18 <- ggplot(merge10, aes(x = L.GBR.GDP, y = FTSE.Close.change)) + 
  geom_point() + 
  geom_smooth() +
  ggtitle("Great Britain") +
  xlab("GDP Growth") +
  ylab("FTSE Growth Rate in %") +
  theme(axis.text=element_text(size=8),
        axis.title=element_text(size=8)) +
  theme(plot.title = element_text(size=10)) 
```

```{r, echo=FALSE, message=FALSE, include=TRUE, cache=FALSE, warning=FALSE}
multiplot(p15, p16, p17, p18, cols = 2)
```

# Inferential Statistics

```{r, echo=FALSE, message=FALSE, include=FALSE, cache=FALSE}
summary(M1 <- lm(DAX.Close.change ~ L.DEU.GDP, data = merge10))
summary(M2 <- lm(NIK.Close.change ~ L.JPN.GDP, data = merge10))
summary(M3 <- lm(CAC.Close.change ~ L.FRA.GDP, data = merge10))
summary(M4 <- lm(FTSE.Close.change ~ L.GBR.GDP, data = merge10))
```

```{r, echo=FALSE, results="asis"}
stargazer(M1, M2, M3, M4,
          type="latex",
          title = "Basic OLS Model",
          dep.var.labels=c("DAX","Nikkei", "CAC", "FTSE"),
          covariate.labels=c("GDP Growth"),
          header = FALSE,
          no.space = TRUE)
```

The above models show us the realtionship between national GDP and the respective stock indices. The predictors are lagged by one annual quarter to mitigate potential endogeneity. All coefficient have a positive sign but are statistically insignificant at the 5 per cent level which suggest that there is no effect differnet from zero. Only the coefficient for French GDP growth is significant at the 5 per cent level. The y-intercepts for all four regressions are positive and statistically significant at the 1 per cent level.

### Heteroscedasticity Diagnose 

```{r, echo=FALSE}
par(mfrow=c(2,2))
plot(M1, which = 1)
  title("Germany")
plot(M2, which = 1)
  title("Japan")
plot(M3, which = 1)
  title("France")
plot(M4, which = 1)
  title("Britain")
```

Looking at the four graphs we see that the basic models suffer to a certain degree from heteroskedasticity. This stems from outliers which bias the results in the graphs. We conclude that the overall degree of hetereoskedasticity is acceptable since the general relationship seems to be random.

### Diagnose of non-normality of Errors

```{r, echo=FALSE}
par(mfrow=c(2,2))
plot(M1, which = 2)
  title("Germany")
plot(M2, which = 2)
  title("Japan")
plot(M3, which = 2)
  title("France")
plot(M4, which = 2)
  title("Britain")
```

The graphs give and idea of the distribution of the error term in the regression. Ideally, the error terms should be normally distributed, and not show a right or left skew. In these graphs, however, we show a cummulative distribution function. That means that we want small residuals in the lower quantiles of the distriution and large residuals in the upper quantiles of the distribution of the error term. Put differently, we'd expect a crosswise line from the lower left bottom to the upper right corner. All four lines sugest such a crosswise pattern indicating the degree of normality of error terms.

```{r, echo=FALSE, message=FALSE, include=FALSE, cache=FALSE}
summary(M5 <- lm(DAX.Close.change ~ L.DEU.GDP + L.DEU.unempl + L.DEU.prvconsm + L.ECB.MRO.change + L.ECB.dep.change + L.WTI.dollar.change + L.Brent.dollar.change, data = merge10))
  
summary(M6 <- lm(NIK.Close.change ~ L.JPN.GDP + L.JPN.unempl + L.JPN.prvconsm + L.ECB.MRO.change + L.ECB.dep.change + L.WTI.dollar.change + L.Brent.dollar.change, data = merge10))

summary(M7 <- lm(CAC.Close.change ~ L.FRA.GDP + L.FRA.unempl + L.FRA.prvconsm + L.ECB.MRO.change + L.ECB.dep.change + L.WTI.dollar.change + L.Brent.dollar.change, data = merge10))

summary(M8 <- lm(FTSE.Close.change ~ L.GBR.GDP + L.GBR.unempl + L.GBR.prvconsm + L.ECB.MRO.change + L.ECB.dep.change + L.WTI.dollar.change + L.Brent.dollar.change, data = merge10))
```

```{r, echo=FALSE, results="asis"}
stargazer(M5, M6, M7, M8,
          type="latex",
          title = "Expanded OLS Model",
          dep.var.labels=c("DAX","Nikkei", "CAC", "FTSE"),
          header = FALSE,
          no.space = TRUE)
```

All coefficients of lagged GDP growth exhibit a positive sign suggesting that an increase in an economy's output leads to a better performance at the stock market. The effect of the coefficients, however, is not significantly different from zero. For unemployment the coefficients are negative for Germany and Japan yet only for the former country the effect is statisitcally significant at the 1 per cent level. Unemployment in France and Great Britain seems to improve the respective stock indices. Again, both effects are not statistically significant, thus no substantial effect can be established.

The pattern that coefficients of Germany and Japan have the same sign as compared to France and Great Britain, holds for private consumption too. For Germany and Japan an increase in private consumption decreases stock indices performances. This effect, however, is not statistically significant in both countries. In France and Great Britain signs are positive which suggests that more spending increases the stock market performance of forms relfected in the indices. This effect is only statistically significant for France (at the 5 per cent level).

Non of the coeffcients of the remaining variables are statistically significant. The signs are for all countries positive and only for Brent price changes negative. The exceptions are Great Britain where increasing change of MRO suggests a decreasing effect on the FTSE, and France where an increasing change in the ECB's deposit facility would hurt the CAC40.

The explanatory power of the models as measured by the Adj.-R^2 is highest for France and Germany, and gets worse for Great Britain and Japan. 

## Heteroscedasticity Diagnose

```{r, echo=FALSE}
par(mfrow=c(2,2))
plot(M5, which = 1)
title("Germany")
plot(M6, which = 1)
title("Japan")
plot(M7, which = 1)
title("France")
plot(M8, which = 1)
title("Britain")
```

Again, the graphs do not suggest large problem with heteroskedasticity. Nonetheless, it must be kept in mind that heteroskedasticity potentially biases the coeffcients systematically. A robust standard error might adjust statistical significance levels by increasing the s.e.-values but leaves the bias untouched.

## Diagnose of non-normality of Errors

```{r, echo=FALSE}
par(mfrow=c(2,2))
plot(M5, which = 2)
title("Germany")
plot(M6, which = 2)
title("Japan")
plot(M7, which = 2)
title("France")
plot(M8, which = 2)
title("Britain")
```

The plotted graph suggests that the standardized residuals lie along the crosswise line. The distribution of the error term is not perfectly normal but is deemed acceptable.

```{r, echo=FALSE, message=FALSE, include=FALSE, cache=FALSE}

summary(P1 <- lm(DAX.Close.change ~ L.DEU.GDP + L.DEU.unempl + L.DEU.prvconsm + L.ECB.MRO.change + L.ECB.dep.change  + L.WTI.dollar.change + L.Brent.dollar.change + year, data = merge10))

summary(P2 <- lm(NIK.Close.change ~ L.JPN.GDP + L.JPN.unempl + L.JPN.prvconsm + L.ECB.MRO.change + L.ECB.dep.change  + L.WTI.dollar.change + L.Brent.dollar.change + year, data = merge10))

summary(P3 <- lm(CAC.Close.change ~ L.FRA.GDP + L.FRA.unempl + L.FRA.prvconsm + L.ECB.MRO.change + L.ECB.dep.change  + L.WTI.dollar.change + L.Brent.dollar.change + year, data = merge10))

summary(P4 <- lm(FTSE.Close.change ~ L.GBR.GDP + L.GBR.unempl + L.GBR.prvconsm + L.ECB.MRO.change + L.ECB.dep.change  + L.WTI.dollar.change + L.Brent.dollar.change + year, data = merge10))
```

```{r, echo=FALSE, results="asis"}
stargazer(P1, P2, P3, P4,
          type="latex",
          title = "Expanded OLS Model with Year Dummies",
          dep.var.labels=c("DAX","Nikkei", "CAC", "FTSE"),
          header = FALSE,
          no.space = TRUE)
```

Taking into account year dummies, the effects change in sign or signifcance or both. In Germany, unemployment becomes statistically insignificant, whereas consumption becomes significant at the 5 per cent level. In Japan, unemployment becomes a positive sign yet the effect is remains statistically insignificant. For France, GDP growth, unemployment, and consumption become a negative sign. Consumption looses its statisticall significance.
GDP growth becomes negative for Great Britain too.

The change in MRO becomes now positive for all coefficients and statistically significant for Germany at the 5 per cent level. The change in the deposit facility is negative for Germany and Japan but positive for the other two countries and throughout not statistically signifcant. Increasing percenatage changes in the WTI price suggest a negative effect on the DAX, the CAC and the FTSE. Only the NIKKEI seems to profit. No effect is statistcally significant. For Brent the signs are reversed.

The years have negative effect throughout. Only for Germany and France the year 2007 has a positive effect. For Germany the year effect becomes positive from 2012 onwards.

The Adj.-R2 is higher for all models than previously. All models have a Adj.-R^2 above 0.9.

## Confidence Intervals

```{r, echo=FALSE}
confint(P1)
confint(P2)
confint(P3)
confint(P4)
```

## Heteroscedasticity Diagnose

```{r, echo=FALSE}
#par(mfrow=c(2,2))

#plot(P1, which = 1)
#title("Germany")

#plot(P2, which = 1)
#title("Japan")

#plot(P3, which = 1)
#title("France")

#plot(P4, which = 1)
#title("Britain")
```

## Diagnose of non-normality of Errors

```{r, echo=FALSE}
#par(mfrow=c(2,2))

#plot(P1, which = 2)
#title("Germany")

#plot(P2, which = 2)
#title("Japan")

#plot(P3, which = 2)
#title("France")

#plot(P4, which = 2)
#title("Britain")
```

```{r, echo=FALSE, message=FALSE, include=FALSE, cache=FALSE}
M13 <- lm(DAX.Close.change ~ L.DEU.GDP + L.DEU.unempl + L.DEU.prvconsm + L.ECB.MRO.change + L.ECB.dep.change  + L.WTI.dollar.change + L.Brent.dollar.change + L.USA.GDP, data = merge10)

M14 <- lm(NIK.Close.change ~ L.JPN.GDP + L.JPN.unempl + L.JPN.prvconsm + L.ECB.MRO.change + L.ECB.dep.change  + L.WTI.dollar.change + L.Brent.dollar.change + L.USA.GDP, data = merge10)

M15 <- lm(CAC.Close.change ~ L.FRA.GDP + L.FRA.unempl + L.FRA.prvconsm + L.ECB.MRO.change + L.ECB.dep.change  + L.WTI.dollar.change + L.Brent.dollar.change + L.USA.GDP, data = merge10)

M16 <- lm(FTSE.Close.change ~ L.GBR.GDP + L.GBR.unempl + L.GBR.prvconsm + L.ECB.MRO.change + L.ECB.dep.change  + L.WTI.dollar.change + L.Brent.dollar.change + L.USA.GDP, data = merge10)
```

```{r, echo=FALSE, results="asis"}
stargazer(M13, M14, M15, M16,
          type="latex",
          title = "Expanded Model with US GDP Growth",
          dep.var.labels=c("DAX","Nikkei", "CAC", "FTSE"),
          header = FALSE,
          no.space = TRUE)
```

```{r, echo=FALSE, message=FALSE, include=FALSE, cache=FALSE}
M9 <- lm(DAX.Close.change ~ L.DEU.GDP + L.DEU.unempl + L.DEU.prvconsm + L.ECB.MRO.change + L.ECB.dep.change  + L.WTI.dollar.change + L.Brent.dollar.change + year + L.USA.GDP, data = merge10)

M10 <- lm(NIK.Close.change ~ L.JPN.GDP + L.JPN.unempl + L.JPN.prvconsm + L.ECB.MRO.change + L.ECB.dep.change  + L.WTI.dollar.change + L.Brent.dollar.change + year + L.USA.GDP, data = merge10)

M11 <- lm(CAC.Close.change ~ L.FRA.GDP + L.FRA.unempl + L.FRA.prvconsm + L.ECB.MRO.change + L.ECB.dep.change  + L.WTI.dollar.change + L.Brent.dollar.change + year + L.USA.GDP, data = merge10)

M12 <- lm(FTSE.Close.change ~ L.GBR.GDP + L.GBR.unempl + L.GBR.prvconsm + L.ECB.MRO.change + L.ECB.dep.change  + L.WTI.dollar.change + L.Brent.dollar.change + year + L.USA.GDP, data = merge10)
```

```{r, echo=FALSE, results="asis"}
stargazer(M9, M10, M11, M12,
          type="latex",
          title = "Expanded Model with Years and US GDP Growth",
          dep.var.labels=c("DAX","Nikkei", "CAC", "FTSE"),
          header = FALSE,
          no.space = TRUE)
```

```{r, echo=FALSE, results="asis"}
vif(M9)
vif(M10)
vif(M11)
vif(M12)
```

Unemployment for Japan becomes negative and stays statistically insignificant. The effect of unemployment in Japan becomes negative whereas signifcance levels do not change. The WTI coefficient for Japan becomes negative. The coefficient for GDP growth in the USA is statisitcall insignificant and positive. Its effect is negative only for France. The Adj.-R2 is again above 0.9 for all models suggesting that more than 90 per cent of variation in the outcome variable can be explained by the model.

## Confidence Intervals

```{r, echo=FALSE}
confint(M9)
```

# Appendix

## Figure X: *Level of Dependent Variables (National Stock Indices) over time*

```{r, echo=FALSE, message=FALSE, include=FALSE, cache=FALSE}
### Germany
Ap1 <- ggplot(data = merge10, aes(x = Date, y = DAX.Close)) + geom_line() + scale_x_continuous(breaks=seq(1991.1, 2015.4, 10)) + ggtitle("Germany | DAX") + theme(plot.title = element_text(size=10)) + labs(x = "Quarters", y = "Av. Closing Value") + theme(legend.position="none") + theme(axis.text=element_text(size=8),
        axis.title=element_text(size=8))
### Japan
Ap2 <- ggplot(data = merge10, aes(x = Date, y = NIK.Close)) + geom_line() + scale_x_continuous(breaks=seq(1991.1, 2015.4, 10)) + ggtitle("Japan | NIKKEI") + theme(plot.title = element_text(size=10)) + labs(x = "Quarters", y = "Av. Closing Value") + theme(legend.position="none") + theme(axis.text=element_text(size=8),
        axis.title=element_text(size=8))
### France
Ap3 <- ggplot(data = merge10, aes(x = Date, y = CAC.Close)) + geom_line() + scale_x_continuous(breaks=seq(1991.1, 2015.4, 10)) + ggtitle("France | CAC40") + theme(plot.title = element_text(size=10)) + labs(x = "Quarters", y = "Av. Closing Value") + theme(legend.position="none") + theme(axis.text=element_text(size=8),
        axis.title=element_text(size=8))
### Great Britain
Ap4 <- ggplot(data = merge10, aes(x = Date, y = FTSE.Close)) + geom_line() + scale_x_continuous(breaks=seq(1991.1, 2015.4, 10)) + ggtitle("Great Britain | FTSE") + theme(plot.title = element_text(size=10)) + labs(x = "Quarters", y = "Av. Closing Value") + theme(legend.position="none") + theme(axis.text=element_text(size=8),
        axis.title=element_text(size=8))
```

```{r, echo=FALSE}
multiplot(Ap1, Ap2, Ap3, Ap4, cols = 2)
```



## Figure X: *Expanded Model: Pooled OLS w/ other Dependent Variable (Levels instead of Change)*

```{r, echo=FALSE, message=FALSE, include=FALSE, cache=FALSE}
summary(AM5 <- lm(DAX.Close ~ L.DEU.GDP + L.DEU.unempl + L.DEU.prvconsm + L.ECB.MRO.change + L.ECB.dep.change + L.WTI.dollar.change + L.Brent.dollar.change, data = merge10))
  
summary(AM6 <- lm(NIK.Close ~ L.JPN.GDP + L.JPN.unempl + L.JPN.prvconsm + L.ECB.MRO.change + L.ECB.dep.change + L.WTI.dollar.change + L.Brent.dollar.change, data = merge10))

summary(AM7 <- lm(CAC.Close ~ L.FRA.GDP + L.FRA.unempl + L.FRA.prvconsm + L.ECB.MRO.change + L.ECB.dep.change + L.WTI.dollar.change + L.Brent.dollar.change, data = merge10))

summary(AM8 <- lm(FTSE.Close ~ L.GBR.GDP + L.GBR.unempl + L.GBR.prvconsm + L.ECB.MRO.change + L.ECB.dep.change + L.WTI.dollar.change + L.Brent.dollar.change, data = merge10))
```

```{r, echo=FALSE, results="asis"}
stargazer(AM5, AM6, AM7, AM8,
          type="latex",
          dep.var.labels=c("DAX","Nikkei", "CAC", "FTSE"),
          header = FALSE,
          no.space = TRUE)
```


## Discussion of Results

Discussion of results
When we look at the negative effect of the change in the deposit facility interest rate in the previous quarter on the stock indices growth rate in the current quarter we see that this effect is always statistically significant at least at the 5 percent. When controlling for year effects, the variable “L.ECB.dep.change” loses its statistical significance for Great Britain. Adding US GDP growth rates does not change this relationship. The negative effect stays insignificant for Great Britain and significant for the other countries.
This seems all the more interesting since the significance of the effect stays for Japan. Great Britain is part of the European Union and one of the Euro Zones most important trading partners therefore, economic interdependencies should be examined and compared between Euro Zone and the two countries. The negative effect suggests that increasing volumes of money held at banks does hurt the stock market, since less money is in circulation which dampens investments.
The effect for previous quarter GDP growth on current stock indices growth is positive which indicates that upward trending stock markets presuppose a healthy economy. This effect is however only statistically significant for Great Britain. In all other countries results do not indicate whether this effect is statistically different from zero.
One possible explanation is that in Great Britain the economy consists to a large extent of the financial sector. Thus, revenue flows from this sector are an important indicator of the economic health in general. In the other countries a well performing economy does not necessarily influence the stock market. One example would be Bosch in Germany, one of the world’s largest firms, which is not listed on the stock market. In conclusion though, one has to take results for Great Britain with caution: if GDP and stock market performance of its firms are intertwined then the regression measures a reverse causality or even a tautology. Put differently, both variables could potentially measure the same concept and the lag would be the only variance in the regression.

